An Elastica-driven Digital Curve Evolution Model for Image Segmentation\thanks{This work has been partially funded by CoMeDiC ANR-15-CE40-0006 research grant.}

### 2019-07-08 jmiv.tex
fatigue

One of the first successful uses of geometric, squared curvature in
image processing is the inpainting algorithm described in
\cite{masnou98inpainting}, where authors evaluate the absolute
curvature along the level lines of a simply-connected image to
reconstruct its occluded parts. The non-intersection property of level
lines induces the construction of an efficient dynamic programming
algorithm. The curvature energy is nevertheless only coarsely
approximated. We may quote \cite{chan02elasticainpainting} as another
geometric inpainting method involving elastica, which is optimized with
level-sets.

In image segmentation, Zehiry and Grady have shown that injecting
curvature as a regularizer can help recover thin and elongated objects
\cite{zehiry10fast}. Similarly Schoenemann {\em et al.}
\cite{schoenemann09linear} have considered the elastica energy in
their ratio-based image segmentation
approach. In~\cite{schoenemann2011elastic}, Schoenemann~{\em et al.}
extend ratio-cut approaches to segmentation with a curvature
term. Their global optimization framework shows the considerable
advantages of using such regularizer in common binary image
segmentation tasks.  However the time complexity of their algorithm
is prohibitive, even if curvature is again coarsely
approximated.  In a more recent work, Lim {\em et al.} propose an
edge-weighted Willmore-like flow to segment
vertebra~\cite{lim2012introducing} using a level-set
formulation. They observe improved results compared
to~\cite{casseles97geodesic} and other perimeter-based priors, but
observe that a shape prior and a close-enough initialization must also be
provided to achieve good results.

@@

The literature presents several perimeter estimators that are multigrid convergent (see
\cite{coeurjolly04comparative,coeurjolly12multigrid} for a review), but in order to define the digital elastica we need a local estimation
of length and we wish that integration over these local length elements gives a multigrid convergent estimator for the
perimeter.

@@

\begin{definition}	
  Given a digital shape $D \subset (h \mathbb{Z})^2$,  the area estimatir $\widehat{Area}(D,h)$ is defined as	
  \begin{align}
    \widehat{Area}(D,h) \coloneqq h^2\text{Card}\left( D \right).	
    \label{eq:digital_estimator_area}
  \end{align}
\end{definition}
It is well known that this area estimator is multigrid convergent
(e.g. see \cite{klette2000multigrid}).  In
\cite{coeurjolly13integral}, the authors combine the
approximation\eqref{eq:curvature_approximation} and digital estimator
\eqref{eq:digital_estimator_area} to define a multigrid convergent
estimator for the curvature.

\begin{definition}[Integral Invariant Curvature Estimator]
  Let $D \subset (h \mathbb{Z})^2$ a digital shape. The integral invariant curvature estimator is defined as
  \begin{align*}
    \hat{\kappa}_{r}(D,x,h) \coloneqq \frac{3}{r^3} \left( \frac{\pi r^2}{2} - \widehat{Area} \left( B_{r} ( x ) \cap D, h \right) \right).
    %% \hat{\kappa}_{r}(D,x,h) \coloneqq \frac{3}{r^3} \left( \frac{\pi r^2}{2} - \widehat{Area} \left( B_{r/h} ( \frac{1}{h} \cdot x ) \cap D, h \right) \right).
  \end{align*}
\end{definition}

@@

In the remaining of this paper we are going to be interested in the minimization of the elastica energy. Given a Euclidean shape $X$ with smooth enough boundary, the elastica is defined as
\begin{align}
  E(X) = \int_{\partial X}{(\alpha + \beta \kappa^2) ds}, \quad \text{for~} \alpha \ge 0, \beta \ge 0.
  \label{eq:elastica}
\end{align}

The digital version of elastica energy approaching the continuous elastica energy is defined as follows (it uses multigrid convergent estimators):
\begin{align}
	\hat{E}( D_h(X) ) = \sum_{\vec{e} \in \partial D_h(X)}{ \hat{s}(\vec{e})\left(\; \alpha + \beta \hat{\kappa}_{r}^2(D_h(X),\dot{\vec{e}},h) \; \right)},
	\label{eq:digital-energy}
\end{align}
where $\dot{\vec{e}}$ denotes the center of the edge $\vec{e}$. In the
expression above, we will substitute an arbitrary subset $Z$ of
$\mathbb{Z}^2$ to $D_h(X)$ since the continuous shape $X$ is unknown.
In the following we omit the grid step $h$ to simplify expressions
(or, putting it differently, we assume that the shape of interest is
rescaled by $1/h$ and we set $h=1$).

@@

The objectives of this section is to show both the potential of minimizing a purely digital elastica energy but also to underline the difficulties of using it in a global combinatorial optimization framework.

Given a digital shape $S^{(0)}$ we describe a process that generates a
sequence $S^{(i)}$ of shapes with non-increasing elastica energy. The
idea is to define a neighborhood of shapes $\mathcal{N}^{(i)}$ to the
shape $S^{(i)}$ and choose the element of $\mathcal{N}^{(i)}$ with
lowest energy.  The process is suited for the integral invariant
estimator but also for other curvature estimators, for example, MDCA
\cite{roussillon11mdca}. As a matter of fact, our experiments have
shown that either estimators induce similar results.

@@

The running time of algorithm \ref{alg:local-search} is summarized in
table \ref{tab:summary-local-comb-rtime}. All the experiments in this paper were executed on a five core $3.4Ghz$ cpu and the number of pixels in the triangle, square and flower shapes are respectively $841,1867,521$ for grid step $h=1.0$. Although its use in practical applications is limited, we
demonstrate that digital estimators are effective in their measurements
and the flows evolve as expected. We observe that it is a
complete digital approach, and we do not suffer from discretization
and rounding problems, a common issue in continuous models.
Furthermore we have checked that this approach works
indifferently with Integral Invariant curvature estimator and
Maximal Digital Circular Arc curvature estimator. So the convergence
of the digital curvature estimator seems to be the cornerstone to
get a digital curve behaving like a continuous elastica.  In the
next section, we explore a more efficient approach.

@@

Recall that the integral invariant estimator approaches curvature by
computing the difference between half of the area of a chosen ball and
the area of the intersection of this ball with the shape.  In
particular, regions of positive curvature have fewer pixels in their
intersection set than on its complement wrt the estimation ball. This
implies that variables in such regions are labeled one, as the
unbalance grows otherwise. We attenuate curvature if we shift the
center of the estimation ball towards the interior of the shape, which
means removing the labeled one pixels. That is why we take the
complement of the optimization solution.

@@

\daniel{In our experiments, the best results are obtained by executing DCE
algorithm with $m$ equal to $r$, where $r$ is the estimation ball
radius (see figure \ref{fig:mx-square-flow} ). We observe that elastica may increase after some iterations if chosen radius is too large, as in the case of the triangle in figure \ref{fig:mx-flow-gs-radius-effect} in which the flow converges to a single point. We conjecture that an appropriated value for the radius should be given by the shape reach.
	The produced flow has no difficulties in handling changes on topology, and it presents different speeds for regions with low and high curvature values, as illustrated in figure \ref{fig:mx-speed-variation-hole-filling}.}

@@

We have studied in depth several digital curve evolution models based
on the elastica energy and we have presented an application of one of
them to image segmentation. The processes we have described are
completely digital, and do not suffer from issues that typically
arises in models that passes through a discretization stage, as
rounding. Moreover, the model can handle changes in topology and its
results are competitive with similar approaches while achieving
reasonable running times.

>>>>

One of the first successful uses of geometric, squared curvature in
image processing is the inpainting algorithm described in
\cite{masnou98inpainting}, where authors evaluate the absolute
curvature along the level lines of a simply-connected image to
reconstruct its occluded parts. The non-intersection property of level
lines induces the construction of an efficient dynamic programming
algorithm. The curvature energy is nevertheless only coarsely
approximated. We may quote \cite{chan02elasticainpainting} as another
geometric inpainting method involving Elastica, which is optimized with
level-sets.

In image segmentation, Zehiry and Grady have shown that injecting
curvature as a regularizer can help recover thin and elongated objects
\cite{zehiry10fast}. Similarly Schoenemann {\em et al.}
\cite{schoenemann09linear} have considered the Elastica energy in
their ratio-based image segmentation
approach. In~\cite{schoenemann2011elastic}, Schoenemann~{\em et al.}
extend ratio-cut approaches to segmentation with a curvature
term. Their global optimization framework shows the considerable
advantages of using such regularizer in common binary image
segmentation tasks.  However the time complexity of their algorithm
is prohibitive, even if curvature is again coarsely
approximated.  In a more recent work, Lim {\em et al.} propose an
edge-weighted Willmore-like flow to segment
vertebra~\cite{lim2012introducing} using a level-set
formulation. They observe improved results compared
to~\cite{casseles97geodesic} and other perimeter-based priors, but
observe that a shape prior and a close-enough initialization must also be
provided to achieve good results.

@@

The literature presents several perimeter estimators that are multigrid convergent (see
\cite{coeurjolly04comparative,coeurjolly12multigrid} for a review), but in order to define the digital Elastica we need a local estimation
of length and we wish that integration over these local length elements gives a multigrid convergent estimator for the
perimeter.

@@

\begin{definition}	
  Given a digital shape $D \subset (h \mathbb{Z})^2$,  the area estimatir $\widehat{\text{Area}}(D,h)$ is defined as	
  \begin{align}
    \widehat{\text{Area}}(D,h) \coloneqq h^2\text{Card}\left( D \right).	
    \label{eq:digital_estimator_area}
  \end{align}
\end{definition}
It is well known that this area estimator is multigrid convergent
(e.g. see \cite{klette2000multigrid}).  In
\cite{coeurjolly13integral}, the authors combine the
approximation\eqref{eq:curvature_approximation} and digital estimator
\eqref{eq:digital_estimator_area} to define a multigrid convergent
estimator for the curvature.

\begin{definition}[Integral Invariant Curvature Estimator]
  Let $D \subset (h \mathbb{Z})^2$ a digital shape. The integral invariant curvature estimator is defined as
  \begin{align*}
    \hat{\kappa}_{r}(D,x,h) \coloneqq \frac{3}{r^3} \left( \frac{\pi r^2}{2} - \widehat{\text{Area}} \left( B_{r} ( x ) \cap D, h \right) \right).
    %% \hat{\kappa}_{r}(D,x,h) \coloneqq \frac{3}{r^3} \left( \frac{\pi r^2}{2} - \widehat{Area} \left( B_{r/h} ( \frac{1}{h} \cdot x ) \cap D, h \right) \right).
  \end{align*}
\end{definition}

@@

In the remaining of this article, we are considering in the minimization of the Elastica energy. Given a Euclidean shape $X$ with smooth enough boundary, the Elastica is defined as
\begin{align}
  E(X) = \int_{\partial X}{(\alpha + \beta \kappa^2) ds}, \quad \text{for~} \alpha \ge 0, \beta \ge 0.
  \label{eq:elastica}
\end{align}

The digital version of Elastica energy approaching the continuous Elastica energy is defined as follows (it uses multigrid convergent estimators):
\begin{align}
	\hat{E}( D_h(X) ) = \sum_{\vec{e} \in \partial D_h(X)}{ \hat{s}(\vec{e})\left(\; \alpha + \beta \hat{\kappa}_{r}^2(D_h(X),\dot{\vec{e}},h) \; \right)},
	\label{eq:digital-energy}
\end{align}
where $\dot{\vec{e}}$ denotes the center of the edge $\vec{e}$. In the
expression above, we will substitute an arbitrary subset $Z$ of
$\mathbb{Z}^2$ to $D_h(X)$ since the continuous shape $X$ is unknown.
In the following we omit the grid step $h$ to simplify expressions
(or, putting it differently, we assume that the shape of interest is
rescaled by $1/h$ and we set $h=1$).

@@

The objectives of this section is to show both the potential of minimizing a purely digital Elastica energy but also to underline the difficulties of using it in a global combinatorial optimization framework.

Given a digital shape $S^{(0)}$ we describe a process that generates a
sequence $S^{(i)}$ of shapes with non-increasing Elastica energy. The
idea is to define a neighborhood of shapes $\mathcal{N}^{(i)}$ to the
shape $S^{(i)}$ and choose the element of $\mathcal{N}^{(i)}$ with
lowest energy.  The process is suited for the integral invariant
estimator but also for other curvature estimators, for example, MDCA
\cite{roussillon11mdca}. As a matter of fact, our experiments have
shown that either estimators induce similar results.

@@

The running time of algorithm \ref{alg:local-search} is summarized in table \ref{tab:summary-local-comb-rtime}. All the
experiments in this paper were executed on a five-core $3.4Ghz$ CPU and the number of pixels in the triangle, square and
flower shapes are respectively $841,1867,521$ for grid step $h=1.0$. Although its use in practical applications is
limited, we demonstrate that digital estimators are effective in their measurements and the flows evolve as expected. We
observe that it is a complete digital approach, and we do not suffer from discretization and rounding problems, a common
issue in continuous models.  Furthermore we have checked that this approach works indifferently with Integral Invariant
curvature estimator and Maximal Digital Circular Arc curvature estimator. So the convergence of the digital curvature
estimator seems to be the cornerstone to get a digital curve behaving like a continuous Elastica.  In the next section,
we explore a more efficient approach.

@@

Recall that the integral invariant estimator approaches curvature by computing the difference between half of the area
of a chosen ball and the area of the intersection of this ball with the shape.  In particular, regions of positive
curvature have fewer pixels in their intersection set than on its complement w.r.t the estimation ball. This implies
that variables in such regions are labeled with 1, as the unbalance grows otherwise. We attenuate curvature if we shift
the center of the estimation ball towards the interior of the shape, which means removing the 1-labeled pixels. That is
why we take the complement of the optimization solution.

@@

\daniel{In our experiments, the best results are obtained by executing DCE
algorithm with $m$ equal to $r$, where $r$ is the estimation ball
radius (see figure \ref{fig:mx-square-flow} ). We observe that Elastica may increase after some iterations if chosen radius is too large, as in the case of the triangle in figure \ref{fig:mx-flow-gs-radius-effect} in which the flow converges to a single point. We conjecture that an appropriated value for the radius should be given by the shape reach.
	The produced flow has no difficulties in handling changes on topology, and it presents different speeds for regions with low and high curvature values, as illustrated in figure \ref{fig:mx-speed-variation-hole-filling}.}

@@

We have studied in depth several digital curve evolution models based
on the Elastica energy and we have presented an application of one of
them to image segmentation. The processes we have described are
completely digital, and do not suffer from issues that typically
arises in models that passes through a discretization stage, as
rounding. Moreover, the model can handle changes in topology and its
results are competitive with similar approaches while achieving
reasonable running times.
